---
title: "STATS 485 Unit 1 Version 1 Paper Reproducibility Appendix"
author: "George Zhou"
date: "January 19, 2019"
output:
  html_document: default
  word_document: default
---
#Overview
The purpose of this appendix is to calculate the Agresti-Coull Confidence intervals for upward mobility of the 40 commuting zones in the United States, and find the upward mobility rate of the four geographical units and see if there is a true difference among them. 

#Setup 
We use the necessary packages, and download the data. 
```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(gridExtra)
library(kableExtra)
mydata = read.csv("http://dept.stat.lsa.umich.edu/~bbh/s485/data/mobility3.csv")
```

#Agresti Confidence Interval + Visualization
The purpose of these blocks are to create a function that will calculate Agresti-Coull 95% CI for upward mobility of the 40 commuting zones. We also create a visualiziation of the CI. 
```{r}
k = 1.96
ac_ci = function(phat, n){
 X = round(phat*n)
 Xt=X+(k^2/2)
 nt=n+k^2
 pt = Xt/nt
 lwr = pt -1.96*(pt*(1-pt))^.5*n^(-.5)
 uppr= pt +1.96*(pt*(1-pt))^.5*n^(-.5)
 if(lwr <0){
   lwr = 0
 }
 if(uppr >1){
   uppr = 1
 }
 return(signif(c(lwr, uppr), digits=3))
}


#testcases
ac_ci(1, 10)
ac_ci(0, 20)

#CI for commuting zones 
a_ci=map(1:40, function(i) ac_ci(mydata$p.upmover[i], mydata$n.lowstart[i]))
a_ci = matrix(a_ci%>%unlist(), nrow = 40, byrow=T)
colnames(a_ci)=c("lowerbound", "upperbound")
mydata1 = data.frame(mydata, a_ci)
table1=mydata1%>%select(zone, region, lowerbound, upperbound)
knitr::kable(table1)
```

The function for creating the graphs. 
```{r}
ci_graph =function(reg, color1, color2){
  graph = ggplot(data = mydata1%>%
                   filter(region == reg), aes(x=zone, y = p.upmover)) + 
    geom_errorbar(mapping = aes(ymin = lowerbound, ymax = upperbound, width =0), colour = color1)+
    geom_point(size = 3, colour = color2)+xlab("Zone")+ylab("Upward Mobility")+
    ggtitle(str_c("Upward Mobility in ", str_to_title(reg), " Region"))+theme_minimal()+coord_flip()
  return(graph)
}
west = ci_graph("west", "seagreen3", "seagreen4")
midwest = ci_graph("midwest", "chocolate2", "chocolate3")
northeast= ci_graph("northeast", "skyblue2", "steelblue2")
south = ci_graph("south", "red2", "red3")

grid.arrange(midwest, northeast, south, west, ncol = 2)
```

#Likelihood Ratio Test and Visualization for difference among 4 regions
The purpose of these blocks is to calculate the upward mobility for the 4 geographical regions, visualize upward mobility based on their geographical regions, and conduct a likelihood ratio test to see if there is a difference among the regions. 
```{r}
ggplot(data = mydata1, mapping = aes(y = p.upmover, x = region, alpha = .4, fill = region)) +geom_boxplot()+ scale_fill_manual(values = c("chocolate2",  "skyblue2","red2", "seagreen2"))+theme_minimal()+theme(legend.position = "none")+ggtitle("Upward Mobility based on Geographical Region")+ylab("Upward Mobility")+xlab("Region")
```

```{r}
table2 = mydata1%>%group_by(region)%>%summarize(p = signif(sum(n.lowstart*p.upmover)/sum(n.lowstart), digits = 3), n = sum(n.lowstart))%>%mutate(region = str_to_title(region))
knitr::kable(table2)%>%kable_styling(bootstrap_options = "striped")
```

We use likelihood ratio test to see if the proportions among the regions are the same. 
```{r}
log_lik = function(p) {
 val = sum (dbinom(n_z*p_ind, size=n_z, prob=p, log=T))
 return(val)
}

n_z = mydata%>%group_by(region)%>%summarize(n = sum(n.lowstart))%>%pull(n)
#common proportion 
p_all = sum(mydata$p.upmover*mydata$n.lowstart)/sum(n_z) 
#proportion for each Region
p_ind=mydata1%>%group_by(region)%>%mutate(x = p.upmover*n.lowstart)%>%summarize(phat=sum(x)/sum(n.lowstart), sum(n.lowstart))%>%pull(phat)
#test statistic 
2*(log_lik(p_ind)-log_lik(sum(p_all*n_z)/sum(n_z)))
qchisq(.95, 3)
#pvalue 
pchisq(8.48, 3, lower.tail = F)
```




